---
title: "disattenuation vs. norm standardised regression slope"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, message = FALSE}
knitr::opts_chunk$set(
    message = FALSE,
    warning = TRUE,
    include = TRUE,
    error = TRUE
)

if (!require(groundhog)) install.packages("groundhog")

groundhog::groundhog.library(c("tidyverse", "faux", "furrr"), 
                             date = "2022-06-14") # older date to accommodate older R version on the R cluster



options(digits = 4, scipen = 999)
theme_set(theme_bw(base_size = 12))
```


# Set up 

## simulation parameters
```{r}
Scenarios = expand_grid(
    # norm sample size
    N_norm = c(50, 100, 250, 500, 1000),
    # local sample size
    N = c(25, 100, 175, 250, 500),
    # population correlation
    roh = c(.1, .3, .5, .7),
    # degree of sample selection: none, above average, half an sd above average, an sd above average (stanine scale)
    selection = c(-999, 5, 6, 7),
    # Simulate each scenario 1000 times
    Simulation = 1:1000)

generate_data_and_calculate_r <- function(N_norm, N, roh, selection,
                                        Simulation) {
  
  # Store scenario's parameters
  Parameters <- environment() %>% as.list() %>% as_tibble()
  
  # function for calculating disattenuated r
  r_disattenuated <- function(x, y, sd_x) {
  
  r_obs <- cor(x,y)
  u_x <- sd(x)/sd_x

 r_obs/(u_x*sqrt(((1/u_x^2)-1)*r_obs^2+1))
  }
  
  # function for norm standardising then calculating regression slope

  b_normed <- function(x, y, sd_x, sd_y) {

  N <- length(x)
  normed_x <- x/sd_x
  normed_y <- y/sd_y
  

 (N * sum(normed_x * normed_y) - sum(normed_x) * sum(normed_y)) / (N * sum(normed_x^2) - sum(normed_x)^2)
  }
  
  
  # Generate Dataset

    # N draws from a bivariate normal distribution with means = 0, SDs = 1 and correlations as above
    x_y <- rnorm_multi(n = 10000, vars = 2, mu = c(5, 5), sd = c(2, 2), r = roh, empirical = TRUE)
    
    x_y_norm_sample <- x_y %>% 
    # sample from the overall sample  with N_norm as defined above
      sample_n(N_norm) 
    
    # extract x and y along with their lengths to make sure it is equal to N
    x_norm <- x_y_norm_sample$X1 
    y_norm <- x_y_norm_sample$X2
      
    x_y_local_sample <- x_y %>% 
    # censor the sample, selecting on X to different degrees as defined above
      filter(X1 > selection) %>% 
    # sample from the censored samples with Ns as defined above
      sample_n(N)

    x_local_sample <- x_y_local_sample$X1 
    y_local_sample <- x_y_local_sample$X2
    
  

  # Our three statistical models predict well-being
  results <- tibble(
    sd_x = sd(x_local_sample),
    sd_y = sd(y_local_sample),
  # calculate correlations
    r_raw = cor(x_local_sample, y_local_sample),
    r_disattenuated = r_disattenuated(x_local_sample, y_local_sample, sd(x_norm)),
    r_norm_standardised = b_normed(x_local_sample, y_local_sample, sd(x_norm), sd(y_norm,))
  ) %>% 
    bind_cols(Parameters, .)  %>% 
    mutate(selection = case_when(
      selection == 5 ~ "above average",
      selection == 6 ~ "half an sd above average",
      selection == 7 ~ "an sd above average",
      TRUE ~ "none"
    ))
  
}
```


## run simulations and save results
```{r, eval = FALSE}
# sim_results = Scenarios %>%
#   pmap(generate_data_and_calculate_r, .progress = T) %>%
#   # Combine everything into a data frame
#   bind_rows()

# We ran this once 

plan(multisession)
sim_results <- Scenarios %>%
  furrr::future_pmap(generate_data_and_calculate_r, .progress = T,
                     .options = furrr::furrr_options(seed = 14)) %>%
  # Combine everything into a data frame
  bind_rows()


write_rds(sim_results, "sim_results.rds")

```

## load results and prepare them for plotting
```{r}
sim_results <- readRDS("sim_results.rds")

sim_results_long <- sim_results %>% 
  pivot_longer(starts_with("r_"), names_to = "r_estimator", values_to = "r_estimate") %>% 
  mutate(selection = factor(selection, levels =c ("none", "above average", "half an sd above average", "an sd above average")),
         bias = r_estimate - roh)
```




# Bias plots
## Norm standardised vs. disattenuated vs. raw
### faceted by selection and roh
```{r, fig.width=8, fig.height=9}
sim_results_long  %>% 
  ggplot(aes(x = as.factor(N), y = bias, group = r_estimator, colour = r_estimator))  +
  geom_line(stat = "summary") +
  geom_pointrange(stat = "summary", fatten = 1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "Bias") + 
  facet_grid(rows = vars(selection), cols = vars(roh))
```

### faceted by selection and N_norm
```{r, fig.width=8, fig.height=9}
sim_results_long  %>% 
  ggplot(aes(x = as.factor(N), y = bias, group = r_estimator, colour = r_estimator))  +
  geom_line(stat = "summary") +
  geom_pointrange(stat = "summary", fatten = 1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "Bias") + 
  facet_grid(rows = vars(selection), cols = vars(N_norm))
```


### across selection degrees and rohs
```{r, fig.width=17/2, fig.height=17/4}
sim_results_long  %>% 
  ggplot(aes(x = as.factor(N), y = bias, group = r_estimator, colour = r_estimator))  +
  geom_line(stat = "summary") +
  geom_pointrange(stat = "summary", fatten = 1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "Bias") + 
  facet_grid(cols = vars(N_norm))
```


## Norm standardised vs. disattenuated
### faceted by selection and roh
```{r, fig.width=8, fig.height=9}
sim_results_long  %>% 
  filter(r_estimator != "r_raw") %>% 
  ggplot(aes(x = as.factor(N), y = bias, group = r_estimator, colour = r_estimator))  +
  geom_line(stat = "summary") +
  geom_pointrange(stat = "summary", fatten = 1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "Bias") + 
  facet_grid(rows = vars(selection), cols = vars(roh))
```


### faceted by selection and N_norm
```{r, fig.width=8, fig.height=9}
sim_results_long  %>% 
  filter(r_estimator != "r_raw") %>% 
  ggplot(aes(x = as.factor(N), y = bias, group = r_estimator, colour = r_estimator))  +
  geom_line(stat = "summary") +
  geom_pointrange(stat = "summary", fatten = 1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "Bias") + 
  facet_grid(rows = vars(selection), cols = vars(N_norm))

```

### across selection degrees and rohs
```{r, fig.width=17/2, fig.height=17/4}
sim_results_long  %>% 
  filter(r_estimator != "r_raw") %>% 
  ggplot(aes(x = as.factor(N), y = bias, group = r_estimator, colour = r_estimator))  +
  geom_line(stat = "summary") +
  geom_pointrange(stat = "summary", fatten = 1) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 9)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "Bias") + 
  facet_grid(cols = vars(N_norm))
```



# SE plots
## Norm standardised vs. disattenuated vs. raw
### faceted by selection and roh
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  group_by(N, roh, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = se, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "SE") + 
  facet_grid(rows = vars(selection), cols = vars(roh))
```

### faceted by selection and N_norm
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  group_by(N, N_norm, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = se, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "SE") + 
  facet_grid(rows = vars(selection), cols = vars(N_norm))
```

### across selection degrees and rohs
```{r, fig.width=17/2, fig.height=17/4}
 sim_results_long %>% 
  group_by(N, N_norm, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  arrange(desc(abs(bias))) %>%
  group_by()  %>% 
  ggplot(aes(x = as.factor(N), y = se, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "SE") + 
  facet_grid(cols = vars(N_norm))
```

## Norm standardised vs. disattenuated
### faceted by selection and roh
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  filter(r_estimator != "r_raw") %>%
  group_by(N, roh, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = se, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "SE") + 
  facet_grid(rows = vars(selection), cols = vars(roh))
```

### faceted by selection and N_orm
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  filter(r_estimator != "r_raw") %>%
  group_by(N, N_norm, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = se, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "SE") + 
  facet_grid(rows = vars(selection), cols = vars(N_norm))
```


### across selection degrees and rohs
```{r, fig.width=17/2, fig.height=17/4}
sim_results_long %>% 
  filter(r_estimator != "r_raw") %>%
  group_by(N, N_norm, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = se, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "SE") + 
  facet_grid(cols = vars(N_norm))
```











# MSE plots
## Norm standardised vs. disattenuated vs. raw
### faceted by selection and roh
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  group_by(N, roh, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = MSE, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "MSE") + 
  facet_grid(rows = vars(selection), cols = vars(roh))
```

### faceted by selection and N_norm
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  group_by(N, N_norm, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = MSE, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "MSE") + 
  facet_grid(rows = vars(selection), cols = vars(N_norm))
```

### across selection degrees and rohs
```{r, fig.width=17/2, fig.height=17/4}
 sim_results_long %>% 
  group_by(N, N_norm, r_estimator) %>%
  summarise(
    bias = mean(r_estimate - roh),
    MSE = mean((r_estimate - roh)^2),
    se = sqrt(sum(r_estimate - mean(r_estimate))^2/(n()-1))
  ) %>%
  arrange(desc(abs(bias))) %>%
  group_by()  %>% 
  ggplot(aes(x = as.factor(N), y = MSE, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2", "#e69f00")) +
  labs(x = "N", y = "MSE") + 
  facet_grid(cols = vars(N_norm))
```

## Norm standardised vs. disattenuated
### faceted by selection and roh
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  filter(r_estimator != "r_raw") %>%
  group_by(N, roh, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = MSE, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "MSE") + 
  facet_grid(rows = vars(selection), cols = vars(roh))
```

### faceted by selection and N_orm
```{r, fig.width=8, fig.height=9}
sim_results_long %>% 
  filter(r_estimator != "r_raw") %>%
  group_by(N, N_norm, selection, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = MSE, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "MSE") + 
  facet_grid(rows = vars(selection), cols = vars(N_norm))
```


### across selection degrees and rohs
```{r, fig.width=17/2, fig.height=17/4}
sim_results_long %>% 
  filter(r_estimator != "r_raw") %>%
  group_by(N, N_norm, r_estimator) %>%
  summarise(
    bias = mean(bias),
    se = sqrt(sum((r_estimate - mean(r_estimate))^2) / (n() - 1)),
    MSE = mean((r_estimate - roh)^2)) %>%
  ggplot(aes(x = as.factor(N), y = MSE, group = r_estimator, colour = r_estimator))  +
  geom_line() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.x = element_line(linetype = "dashed", size = 0.3),
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  ) +
  geom_line(y = 0, colour = "black") +
  scale_colour_manual(values = c("#009e73", "#0072b2")) +
  labs(x = "N", y = "MSE") + 
  facet_grid(cols = vars(N_norm))
```











